Privacy and Security Benefits of Jailbreaking iOS

### 2013-12-13 final_project/paper.tex
final sorta commit

\section{Introduction}
\tab The App Store is an integral part of the iOS security model. All applications on the App Store must go through an approval process, one of whose purposes is to prevent malicious apps from reaching users. This process involves quality control, static and manual analysis of the behavior and content of apps. In addition the use of private APIs, which typically would allow an app to provide unapproved or malicious functionality is screened against. Apps on the iOS platform are assigned a unique identifier and a folder whose contents it can modify, while the rest of the filesystem is out of reach. Individual apps can communicate with other apps in a very limited fashion by means of URL Handlers. Access to Apple services such as Push Notifications, Game Center and Passbook is authenticated through certificates issued by Apple. Moreover, every application on the App Store is code signed, and code that hasn't been signed by Apple cannot run on the device. All these measures ensure that apps run in a sandboxed environment which ensures that the system and the user data on the device are protected in case an app attempts malicious activity.

Jailbreaking is the process of modifying iOS through what is most often a security exploit in order to allow code that hasn't been signed by Apple to run on the device, in addition to enabling root access to the filesystem for any app. Hence jailbreaking allows for downloading apps from sources exempt of Apple's stringent approval procedures, which typically enhance the functionality of the OS and its stock apps and allow for unapproved use of private API that provide greater functionality and freedom to developers.

 Below I outline researchers' findings on circumventing the approval process and examples of malicious apps on the App Store, in order to expose the reliance on the walled garden model as insufficient with regards to ensuring security and privacy.

@@

Considering these important uses of jailbreaking, in this paper I present third-party research that benefits from jailbreaking in evaluating the state of security and privacy on iOS, and showcase useful tools that can improve the security and privacy of jailbroken iOS users.

@@

In reality, since the approval process does not require that developers submit their source code for a review, it is quite possible to use code obfuscation techniques in order to evade any static or dynamic code analysis tools Apple might be using to uncover threats.  In his paper on iOS privacy\cite{priv}, Nicolas Seriot demonstrates several simple techniques that can be used to deliberately circumvent common types of code analysis likely used by Apple in search of calls to private APIs and system directories. He suggests that his proof-of-concept SpyPhone app\footnote{https://github.com/nst/spyphone/} in fact does not rely on private APIs in the first place, and goes on to express concern that the existence of apps on the App Store collecting personal data is quite likely given the underground demand for and abundance of personal data on iOS devices.

@@

\cite{attacks}, 
\section{Threats to Privacy in the AppStore}
Even though researchers have demonstrated techniques such as dynamic \cite{dynamic} and static \cite{pios} analysis targeted at capturing data flows within the application have been developed, several controversial cases of user data misuse and user profiling have gained attention\footnote{http://mclov.in/2012/02/08/path-uploads-your-entire-address-book-to-their-servers.html}.

@@

While AppStore apps can pose threats to privacy, iOS itself collects large amounts of data from normal usage.
In 2011, a controversy arose due to iOS maintaining a database of WiFi hotspots and cell towers around the users location, in order to improve its geo-location services beyond complete reliance on the slower GPS systems which would allow for quickly determining the user's current location when using Maps.\footnote{http://www.networkworld.com/community/blog/apple-officially-responds-ios-consolidateddb-} Even though, Apple assured that despite this location data being sent to Apple servers, it is completely anonymous the data would remain on the device itself in a file named consolidated.db which can then be easily retrieved.\footnote{https://www.sans.org/reading-room/whitepapers/forensics/forensic-analysis-ios-devices-34092}.

@@

\section{Threats of Jailbreaking}

you can crack the pin code of someone by jailbreaking an un jailbroken phone first and then changing the pin
you can use the espionage tools in \cite{espionage} unless you have stuff fixed.
default root password is very well known.
can't update if u want to keep the jailbreak.
installing malicious apps from evil repositories.

\section{Action Items}
\subsection{Jailbreaking for better privacy}

Before address book permissions were first added to iOS 6, the well publicized case of the social networking app Path drew attention to the issue of privacy in the AppStore. It was discovered that Path collected personal information from the user's address book and sent it to company servers without the users consent and it did not list this activity in the privacy agreement. The creators of Path have since removed the behavior and settled the ensuing FTC charges. \footnote{http://www.ftc.gov/news-events/press-releases/2013/02/path-social-networking-app-settles-ftc-charges-it-deceived} In response to this, a tweak for jailbroken iOS devices named ContactPrivacy appeared. It allowed the user to control access to the address book.

@@

One such vulnerability\footnote{http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2011-0226} in the PDF rendering engine of Apple iOS\cite{pios}, was successfully used to jailbreak earlier version of iOS by allowing the code necessary to obtain elevated privileges run on the device.

\section{Conclusion}
The problem of course is somewhat intractable, because trusting data to a third party comes with no guarantees about how the data will be used. Nevertheless, it is important to make sure that users are given the power to \emph{decide} whether to give their data in the first place. This privacy enhancement feature can be a great complement to pre-existent code approval policies and jailbreaking has allowed for several such data flow management tools. However it wasn't until iOS7 before users were able to control microphone use or limit ad tracking\footnote{http://blogs.wsj.com/digits/2013/09/18/how-to-use-apples-new-ios-7-privacy-controls/} and a dedicated Privacy Settings configuration pane providing consolidated control over the permissions of individual apps has been available only as of iOS6\footnote{https://developer.apple.com/library/ios/releasenotes/General/RN-iOSSDK-6\_0/}.

%\subsection{}

>>>>

\section{Introduction}
\tab The App Store is an integral part of the iOS security model. All applications on the App Store must go through an approval process, in part to prevent malicious apps from reaching users. This process involves quality control, static and manual analysis of the behavior and content of apps. In addition, the use of private APIs, which typically would allow an app to provide unapproved or malicious functionality is screened against. Each app on the iOS platform is assigned a unique identifier and a folder whose contents it can modify, while the rest of the filesystem is out of reach. Individual apps can communicate with other apps in a very limited fashion by means of URL Handlers. Access to Apple services such as Push Notifications, Game Center and Passbook is authenticated through certificates issued by Apple. Moreover, every application on the App Store is code signed, and code that hasn't been signed by Apple cannot run on the device. All these measures ensure that apps run in a sandboxed environment which ensures that the system and the user data on the device are protected in case an app attempts malicious activity.

Jailbreaking is the process of modifying iOS through what is most often a security exploit in order to allow code that hasn't been signed by Apple to run on the device, in addition to enabling root access to the filesystem for any app, effectively subverting the sandboxing on iOS. Since jailbreaking allows the execution of code that has not been signed by Apple, it makes it possible to download and install apps and tweaks from sources exempt of Apple's stringent approval procedures. Such apps and tweaks typically enhance the functionality of the OS and its stock apps or boast features that are impossible to implement with the official SDK, as jailbreaking permits developers to use officially unapproved private APIs that provide additional functionality.

@@

Considering these important uses of jailbreaking, in the following sections I outline the findings of researchers regarding privacy and security on iOS, alongside recent cases of privacy flaws in iOS and apps from third-party developers. In the section Action Items I showcase useful tools that can improve the security and privacy of jailbroken iOS users and how they pertain to real threats.

@@

In reality, since the approval process does not require that developers submit their source code for a review, it is quite possible to use code obfuscation techniques in order to evade any static or dynamic code analysis tools Apple might be using to uncover threats.  In his paper on iOS privacy\cite{priv}, Nicolas Seriot demonstrates several simple techniques that can be used to deliberately circumvent common types of code analysis likely used by Apple to search for calls to private APIs and system directories. Seriot suggests that his proof-of-concept SpyPhone app\footnote{https://github.com/nst/spyphone/} in fact does not rely on private APIs in the first place, and goes on to express concern that the existence of apps on the App Store collecting personal data is quite likely given the underground demand for and abundance of personal data on iOS devices.

@@

%\cite{attacks}, 
\section{Threats to Privacy in the AppStore}
Even though researchers have demonstrated techniques such as dynamic \cite{dynamic} and static \cite{pios} analysis targeted at capturing data flows within the application have been developed, several controversial cases of user data misuse and user profiling have gained attention\footnote{http://mclov.in/2012/02/08/path-uploads-your-entire-address-book-to-their-servers.html}.

@@

While AppStore apps can pose threats to privacy, iOS itself collects large amounts of data from normal usage.
In 2011, a controversy arose due to iOS maintaining a database of WiFi hotspots and cell towers around the users location in order to improve its geo-location services beyond complete reliance on the slower GPS systems which would allow for quickly determining the user's current location when using Maps.\footnote{http://www.networkworld.com/community/blog/apple-officially-responds-ios-consolidateddb-} Even though Apple assured that the location data is being sent to Apple servers in an anonymized manner, the data would remain on the device itself in a file named consolidated.db which could be easily retrieved.\footnote{https://www.sans.org/reading-room/whitepapers/forensics/forensic-analysis-ios-devices-34092}.

@@

\section{Action Items}
\subsection{Jailbreaking for improved privacy control}

Before address book permissions were first added to iOS 6, the well publicized case of the social networking app Path drew attention to the issue of privacy in the AppStore. It was discovered that Path collected personal information from the user's address book and sent it to company servers without the users consent and it did not list this activity in the privacy agreement. The creators of Path have since removed the behavior and settled the ensuing FTC charges.\footnote{http://www.ftc.gov/news-events/press-releases/2013/02/path-social-networking-app-settles-ftc-charges-it-deceived} In response to this, a tweak for jailbroken iOS devices named ContactPrivacy appeared. It allowed the user to control access to the address book.

As of iOS7 Apple doesn't request the users permission whenever malicious apps request to use the came for capturing video or photos. There are proof of concept spyware apps that allow surreptitious camera captures\cite{espionage} By hooking the methods for requesting the camera, I developed a tweak for MobileSubstrate (the code injection platform on most jailbroken devices). The simple tweak, named CameraSentry, allows the user to deny camera access to malicious spy apps.

\subsection{Changing the default root password}

Users with jailbroken devices are advised to change the default root password - the well known "alpine". Since an attacker can install a remote login utility such as SSH and gain full control over the device, in case a device is jailbroken by an attacker, the default password (if left unchanged) can be used to access the filesystem, bypassing any lock screen passwords the user might be using (passwords are stored in \texttt{{\textbackslash}var{\textbackslash}Keychain})\cite{lostpass}

\subsection{Installing an outbound connections firewall}
Since iOS doesn't come with a user configurable firewall, it is impossible for the end user to monitor the outbound connections that apps attempt. In the case of privacy leaking applications this is of utmost importance as the user can prevent the leaking of information to third party servers shall they notice any suspicious data transfers (e.g. a game connecting to a server different than Apple's Game Center servers. A solution (Firewall iP 2.0) for outbound connection monitoring and control can be downloaded from the Cydia jailbreak app.
\subsection{Disabling anonymous usage statistics}

Developers use code from third-parties to collect usage statistics for their apps just like websites use third party solutions like Google Analytics to obtain insight about their customer base. Installing a jailbreak tweak like PrivaCy\footnote{http://cydia.saurik.com/package/com.saurik.privacy} allows end users to effectively disable sending such statistics.

@@

\section{Conclusion}
While the AppStore is well known for its stringent approval policies, in light of the security research and controversial cases presented in this paper it is clear that stringent approval policies alone are not sufficient to protect users. Better static and dynamic analysis techniques are being researched\cite{dynamic} thanks to the runtime access provided by jailbreaking and can be used to enhance the Apple's screening process. Moreover, since mobile devices are highly personal, they contain valuable personal data which makes any potential security risk a serious privacy risk first and foremost. In this sense security and privacy are tightly related on iOS and the problem of privacy is somewhat intractable even when knowingly trusting data to a third party because there are no guarantees about how the data will be used. However, it is important that platforms evolve towards more power to the users when \emph{deciding} whether to give their data in the first place, complemented by the ability to fine tune the flow of sensitive data. Overall the enhancements that jailbreaking brings provide more control over privacy and can help uncover security threats by allowing unhindered security research on Apple's otherwise source platform.
%\subsection{}
\pagebreak
